A Multi-Modal Model of Object Deformation for Robotic Pushing

### 2016-09-09 twostagedo.tex
many small typos corrected.

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
% Do not put math or special symbols in the title.
% could we change the title to this? What do you think? I wanted to get the modular idea in there.
\title{A Multi-Modal Model of Object Deformation for Robotics
%A Multi-Modal Model for Prediction and Classification of Object Deformation during Robotic Manipulation
}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%

@@

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
\comment{In this paper we present a multi-modal framework for learning generative models of object deformation under robotic pushing. The model is multi-modal in that it is based on integrating force and visual information. The framework consists of several sub-models that are independently learned from the same data. These component models can be sequenced to provide many-step prediction and classification. When presented with a test example--a robot finger pushing a deformable object made of an unidentified, but previously learned, material--the predictions of modules for different materials are compared so as to classify the unknown material.} Our approach, which consists of learning and combining multiple models, goes beyond previous techniques by enabling i) predictions over many steps, ii) learning of plastic and elastic deformation from real data, iii) prediction of forces experienced by the robot, iv) classification of materials from both force and visual data, v) prediction of object behaviour after contact by the robot terminates.  While previous work on deformable object behaviour in robotics has offered one or two of these features none has offered a way to achieve them all, and none has offered classification from a generative model. We do so through separately learned models which can be combined in different ways for different purposes.
\end{abstract}

@@

\subsection{Integration Scheme}
\label{sec:ischeme}
Rather than using the traditional Newton equations with a numeric integration scheme like Euler \comment{or} Verlet, \comment{used twice to solve the second order differential equiations,} we made the \comment{\sout{velocities proportional to the forces:} force proportional to the velocity:
\begin{align}
 F(t) = m \dfrac{\partial \vec{p}}{\partial t}
\end{align}
and applied Euler only once to solve for $p$:
\begin{equation}
 \vec{p}(t+h) = \vec{p}(t) + h \dfrac{\vec{F(t)}}{m}\label{eq:integration}
\end{equation}}
with $\vec{F}(t) = \vec{F}_D(t) + \vec{F}_A(t) + F_\varphi(t)$ being the sum of the forces for the preservation of length, area and angle respectively and $\vec{p}(t+h)$, the position of the vertex at time $t+h$.  \comment{We acknowledge that this is not the proper Newtonian model, however t}he simulations obtained were closer to the observed behaviour, because there is not inertia in the oscillation.
When experimenting with different integration steps, force values were interpolated between measurements.

\subsection{Collision Detection}
In the simulation, rigid objects \comment{(like the robotic finger and the table)} are treated as geometric obstacles.  \comment{When the simulation of the behaviour of the mesh, which is used to model the sponge, is run, the predicted positions of the vertices may indicate that the mesh penetrates the obstacles.  This should be interpreted as a collision between the sponge and the obstacle.  Since the sponge can not penetrate an obstacle, these overlaps must be eliminated. %The physical consequences of this will affect the system predictions.  For example: if the vertices of the mesh overlap the table, the sponge will remain on top of the table, with its material being more compact, which will cause the response force propagated towards the finger to be increased.
}

In order to model this, we model the finger and table using geometric primitives.  For the table obstacle, any interpenetration by \comment{the sponge} is solved by pushing vertices \comment{of the mesh of the sponge} back to the border of the table.  The springs will then propagate the effect to their neighbours.  For the finger obstacle, resolution is in two stages: by vertices and then edges.  A vertex will be pushed out to a distance $\epsilon$ of the finger, in the direction of the radius.  For edges, the standard algorithm in \alref{alg:point_segment} is used to estimate the shortest distance between the centre of the finger and the edge.  Given this distance, the algorithm in \alref{alg:push_edge} indicates how far and in which direction the edge must be displaced.

\begin{figure}
  \begin{algorithmicieee}{Closest Point on a Segment to Any Other Point}\label{alg:point_segment}
 \STATE Be $a$, $b$ the ends of the edge $\overline{AB}$, $c$ the external point, and $x$ the closest point to $c$ in the line.
 \STATE Be the vectors $\vec{AB} = b - a$ and $\vec{AC} = c - a$.
 \STATE The projection of $\vec{AC}$ over $\vec{AB}$ is $\|x\| = \dfrac{\vec{AC} \cdot \vec{AB}}{\|\vec{AB}\|}$
 \IF{$\|x\| < 0$}
  \STATE $closest \gets a$
 \ELSE
  \IF{$\|x\| > \|\vec{AB}\| $}
   \STATE $closest \gets b$
  \ELSE
   \STATE $closest \gets a + \|x\| \cdot \dfrac{\vec{AB}}{\|\vec{AB}\|}$
  \ENDIF
 \ENDIF
 \end{algorithmicieee}
 \caption{Algorithm to find the closest point on a segment to any other point.}
\end{figure}

\begin{figure}
  \begin{algorithmicieee}{Push Edge Out}\label{alg:push_edge}
 \STATE Be $a$, $b$ the ends of the edge $\overline{AB}$, $c$ the centre of the circle.  Use algorithm \ref{alg:point_segment} to find the closest point to the centre in the edge.
 \IF {$\|(closest - c)\| < radius$ (There is an intersection)}
  \STATE $displacement = (radius - \|(closest - c)\| + \epsilon) * \dfrac{(closest - c)}{\|(closest - c)\|}$
  \STATE $ a \gets a + displacement$
  \STATE $ b \gets b + displacement$
 \ENDIF
\end{algorithmicieee}
 \caption{Algorithm to push an edge out of a circular obstacle.}
\end{figure}

\subsection{Geometric Constraints}
Even with the addition of the term for preservation of angles, the discretized mass-spring forces can cause triangles of the mesh to overlap each other and be reversed or to be flattened.  The option of giving additional aid through small geometric subroutines was evaluated as well.  There are two geometric constraints that the mesh is enforced to maintain:

\begin{enumerate}
 \item For a reversed triangle, (one vertex crossed over an edge): the vertex gets pushed beyond half the distance between the vertex and the edge in the direction of the ``reversed'' height\footnote{$3/4$ this distance was used in our implementation.}, while both ends of the edge get pushed the same amount in the opposite direction.  This is enough for most cases, but triangles in the border of the mesh can still overlap each other.

@@

Given the initial shape of the object and a set of parameters, the mass-spring system allows to make predictions of the deformation of an object for various interactions.  Given the procedures explained in the previous sections, there are three ways to calculate the subsequent deformations of the shape:
\begin{enumerate}
 \item \textbf{Collision detection only.} Collision detection is used to determine the compression of the springs around the actuator.  In this case, the neighbouring springs are compressed and their preservation forces propagate the deformation towards their neighbours, and so on.  There is no need to use the force sensor.  It can be equated with observing someone else pushing an object we have touched before.
 \item \textbf{Forces on vertices.} The forces to be applied on the  vertices around the actuator are approximated by the measured reaction forces.  Both, the series of positions of the actuator and the reaction forces are necessary.  These forces will be solely responsible for the deformation of the mesh.  If the deformation is insufficient, the actuator may trespass invalidly the border of the mesh.  The quality of the simulation is much more sensitive to the values of the elasticity-plasticity constants.  This corresponds to the robot itself pushing the object and feeling the response, while using vision solely to determine the position of the actuator and for evaluation purposes.
 \item \textbf{Both.} The previous two are combined.  Visual constraints (collision detection) and forces are used to determine the displacement of nodes around the actuator.  Again, the springs propagate the deformation to the rest of the mesh.
\end{enumerate}
The mesh as predicted by one step is fed as current shape for the next step and the cycle is repeated until the simulation is stopped.

The next stage \commentErrase{of} \comment{is} the learning process \comment{which} consists in identifying sets of parameters that produce simulations similar to the ground truth.  This problem is addressed in the following section.

\section{Learning to Predict the Deformation}
\label{sec:learning}

@@

\begin{figure}[!t]
\centering
\includegraphics[width=88mm]{arrio2}
\caption{(a) Experimental set up.  A robotic finger pushes a block of deformable material over a table.  A camera registers the events of the plane where the main deformations take place.  (b) Evaluation function.  Left top: Mesh over real block of material.  Left bottom: Tipical error of the model vs the ground truth (red) and tipical error of the tracker (blue).  Right top: tracked ground data.  Middle: TP in yellow, FP in blue, FN in red and TN in gray.  Bottom: simulated mesh. }
\label{fig:scene}
\end{figure}

The contour of the deformable object is used to obtain the ground truth for the training phase, which is the area enclosed by the contour. \comment{Since we only require a tracking algorithm which is good enough to evaluate the predictions of the mass-spring system, a very simple one was used.}  Geometrically, the contour is represented by a polygon with hundreds of sides, which can also be called a \textit{linear spline} \comment{\cite{Menet1991}. The vertices of the polygon work as control points which can be moved to adjust the polygon to the shape of the object as it gets deformed.  Even though this implementation is inspired in the theory of active contours, it works with a much simpler and faster, but less accurate algorithm as it is explained below, instead of minimizing a global energy.}  The tracking of the deformable target in the ground truth has two stages:
\begin{enumerate}
 \item The target object is segmented from the environment.  The Canny algorithm is used to extract the edges of the image.  \comment{\sout{Next, a simple colour segmentation is used to remove edges that are not likely to belong to the object of interest.}  Since the object of interest has an outstanding colour, the hue of the pixels on every edge are used to keep only edges that could belong to it.  If the average hue of the pixels on an edge fall outside a previously calibrated range, the edge is eliminated.}
 \item The internal representation of its contour, the linear snake, is adapted to its new shape.  The algorithm assumes that no occlusions take place and that the contour does not bend over itself.
\end{enumerate}
For the first frame, the snake is initialised as a rectangle around the target. This rectangle fragments itself into smaller pieces. Each new vertex adheres itself to the closest point on an edge, as reached by growing rings of pixels around its current position, up to a certain threshold distance [\fref{fig:scene}(a)].  If there is no edge within that area, the control point remains were it is. This criterion was selected to make the representation more robust against edges that are easier to detect on some frames than in others.  Once the edge reappears, the control point adapts itself again.  Also, the distance between two consecutive control points is kept within the interval $[min\_distance, max\_distance]$ and two adjacent edges do not form angles smaller than $min\_angle$.  See \alref{alg:regularise}.  This allows the snake to adapt very quickly to the complexity of the contour it represents.

\begin{figure}
 \begin{algorithmicieee}{Regularise Linear Spline}\label{alg:regularise}
 \STATE Be $L$ a linear spline representing the contour of a $2D$ shape, and $L_i$ the $i^{th}$ control point in a cyclic order.
 \FORALL {$L_i$ in $L$}
 \IF {$dist(L_i,L_{i-1}) > max\_distance$}
  \STATE Insert $midpoint(L_i,L_{i-1})$ after $L_i$
 \ELSIF {$dist(L_i,L_{i-1}) < min\_distance$}
  \STATE erase $L_i$
 \ELSIF {$angle(L_{i-1},L_i,L_{i+1}) < min\_angle$}
  \STATE erase $L_i$
 \ENDIF
 \ENDFOR
\end{algorithmicieee}
\caption{Algorithm used to regularise the number of elements of a linear spline, in accordance with the level of detail required.}
\end{figure}

\subsection{Evolutionary Algorithm}
\comment{Since a mass-spring model is not a constitutive model, its parameters are not directly derived from the physical properties of the material, they depend on the geometry of the mesh used to model the object, and there can be several sets of parameters in different regions of the space of parameters that produce adequate simulations \cite{Morris2008}, we chose to use an evolutionary algorithm to expore different regions and try improvements in the vecinity of successful individuals.}

The evolutionary algorithm listed in \alref{alg:genetic} was used to search the space of parameters of the mass-spring model presented in the previous section.
From the parameters that must be chosen for the simulation to work, some were fixed while others had to be searched for.  They are:
\begin{itemize}
 \item The mass per vertex.  \textit{(For the moment we consider a unitary mass for all vertices.)}
 \item The elasticity constants: for the preservation of length $k_L$, area $k_A$, angles $k_{\varphi}$ and linear damping $k_d$.  \textit{(All springs will have the same values to favour generalisation.)}
 \item The plastic parameters: $yield$, $creep$ and the maximum amount of permanent deformation $max\_\alpha$.  \textit{(All springs will have the same values to favour generalisation.)}
 \item The integration time step $\Delta t$.  \textit{(It has been set to 0.1 seconds, even though values as small as 0.01 were tried as well.)}
 \item An upper limit for the magnitude of the total estimated forces per vertex $max\_Force$.
 \item The minimum area before a triangle is considered \textit{flat}.  (It has been set to $2.0$ units.)
\end{itemize}
To evaluate the suitability of each set of parameters, the simulation is compared with the ground truth tracked from real objects.

\begin{figure}
  \begin{algorithmicieee}{Evolutionary Search}\label{alg:genetic}
 \STATE Be $N$ the number of sets of parameters to be evaluated per generation and $M$ the number of generations to try.
 \STATE Let $\alpha + \beta + \gamma + \zeta = N$ denote four portions of the $N$ elements\comment{, with: \\
 $\alpha \leftarrow$ number of elements that survive for the next generation as they are, \\
 $\beta \leftarrow$ mutated individuals, \\
 $\gamma \leftarrow$ crossed over, and \\
 $\zeta \leftarrow$ new individuals. \\
 which will be chosen as follows:}
 \STATE Be $\sigma_{max} > \sigma_{min}$ standard deviations for a Gaussian
 \STATE Let $\Delta\sigma \gets (\sigma_{max} - \sigma_{min})/M$.
 \FOR {$i=1 \to M$}
 \STATE Run the $N$ simulations and compare prediction with the ground truth (real object), by comparing the area occupied by the real object with the area occupied by the mesh of the simulation [\fref{fig:scene}(b)].
 \STATE Keep the best $\alpha$ of all simulations.
 \STATE $\sigma \gets \sigma - \Delta\sigma$
 \STATE Generate $\beta$ by adding random Gaussian noise, with standard deviation $\sigma$, to the best candidates.
 \STATE Generate $\gamma$ by selecting values of the parameters from two different sets at random, after eliminating the $\epsilon$ worst.
 \FOR {$\zeta$ elements}
   \STATE Generate completely new sample with:
   \FOR {every parameter}
     \STATE $n \leftarrow randomNumber \in [0,1)$
     \STATE $min \leftarrow \log(minValue)$
     \STATE $max \leftarrow \log(maxValue)$
     \STATE $param \leftarrow e^{(n (max - min) + min)}$
   \ENDFOR
 \ENDFOR
 \STATE $i \gets i + 1$
 \ENDFOR
\end{algorithmicieee}
 \caption{Evolutionary Search for Automatic Calibration}
\end{figure}

@@

To evaluate the quality of the predicted shapes, the pixels inside the tracked contour are used as ground truth.  These are  compared against those inside the mesh of the simulation.  In machine learning terms, these pixels can be divided among four classes: the \textit{true positives $TP$}, where the material is present, and the simulation predicted it would be present; the \textit{true negatives $TN$}, where the absence of material is in accordance with the prediction; the \textit{false positives $FP$}, where the prediction said there would be material, but there is not; and the \textit{false negatives $FN$}, where there is material, but the prediction said there wouldn't [\fref{fig:scene}(b)].  Given that maintaining a record of the true negatives (the empty/background space around the material) is mostly irrelevant and highly inefficient, the mark for a whole video was derived from the following quantities:
\begin{align}
 Precision &= \frac{TP}{TP+FP}, \\
 Recall &= \frac{TP}{TP+FN}
\end{align}
From which the harmonic average $F$ is derived:
\begin{align}
 F = \frac{2}{\frac{1}{Precision} + \frac{1}{Recall}} = \frac{1}{1 + \frac{FN + FP}{2TP}}
\end{align}

This measurement \comment{is know as F-score and} is used in the computer vision community to evaluate segmentation algorithms.  $F=1$ means that there were no false predictions ($FN+FP = 0$).  $F \rightarrow 0$\footnote{This is read: the value of $F$ goes to zero or $F$ tends to zero.}, if there are no good predictions ($TP\rightarrow0$) or if there are too many errors ($FN+FP \rightarrow \infty$).  $F$ is calculated for every frame and the mark for the video \comment{ at frame $i$} is the mean over all the frames so far, $\mu(F)$:
\comment{\begin{align}
 \mu(F) = \sum_{t=0}^{t=i} F(frame=t)
\end{align}}
If the predictions for all the frames were perfect, $\mu(F)$ would be equal to 1; the closest to zero, the more errors there were.  The simulation stops sometimes, if a model has become unstable; from that frame on, the value of $F$ is zero.  The model with the highest final value of $\mu(F)$ is considered the best.

\section{Models of Force Learning}
%We used two methods to predict the forces.  The first one consists in measuring the total force acting on the vertex of the mesh closest to the actuator, according to the mass-spring model, after solving detected collisions.  This force is feed as input for the following simulation step.
\label{sec:forces}

Additionally and independently from the mass-spring system,  it is possible to make predictions about the expected reaction force of the material, by using information from the stress-strain diagram obtained from the \comment{same} training data set \comment{than the mass-spring model.  This diagram is obtained from the video recorded by the visual system of the robot and the forces recoded by a sensor located at the wrist.  When the readings of the force sensor start to increase and the figer collides with the block of material, the point of zero deformation is identified.  From here on, the displacement of the finger is considered equivalent to the amount of deformation in the material.  Even though this method is not as rigorous as an engineering test, the estimation is good enough for robot manipulation of common materials.}

>>>>

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
% Do not put math or special symbols in the title.
% could we change the title to this? What do you think? I wanted to get the modular idea in there.
\title{A Multi-Modal Model of Object Deformation for Robotic Pushing
%A Multi-Modal Model for Prediction and Classification of Object Deformation during Robotic Manipulation
}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%

@@

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
\comment{In this paper we present a multi-modal framework for learning generative models of object deformation under robotic pushing. The model is multi-modal in that it is based on integrating force and visual information. The framework consists of several sub-models that are independently learned from the same data. These component models can be sequenced to provide many-step prediction and classification. When presented with a test example--a robot finger pushing a deformable object made of an unidentified, but previously learned, material--the predictions of modules for different materials are compared so as to classify the unknown material.} Our approach, which consists of learning and combining multiple models, goes beyond previous techniques by enabling i) predictions over many steps, ii) learning of plastic and elastic deformation from real data, iii) prediction of forces experienced by the robot, iv) classification of materials from both force and visual data, v) prediction of object behaviour after contact by the robot terminates.  \comment{While previous work on deformable object behaviour in robotics has offered one or two of these features none has offered a way to achieve them all, and none has offered classification from a generative model. We do so through separately learned models which can be combined in different ways for different purposes.}
\end{abstract}

@@

\subsection{Integration Scheme}
\label{sec:ischeme}
\comment{Rather than using traditional Newtonian equations with a numeric integration scheme like Euler or Verlet, used twice to solve the second order differential equiations,} we made the \comment{force proportional to the velocity:
\begin{align}
 F(t) = m \dfrac{\partial \vec{p}}{\partial t}
\end{align}
and applied Euler only once to solve for $p$:
\begin{equation}
 \vec{p}(t+h) = \vec{p}(t) + h \dfrac{\vec{F(t)}}{m}\label{eq:integration}
\end{equation}}
with $\vec{F}(t) = \vec{F}_D(t) + \vec{F}_A(t) + F_\varphi(t)$ being the sum of the forces for the preservation of length, area and angle respectively and $\vec{p}(t+h)$, the position of the vertex at time $t+h$.  \comment{We acknowledge that this is not the proper Newtonian model, however t}he simulations obtained were closer to the observed behaviour, because there is no inertia in the oscillation.
When experimenting with different integration steps, force values were interpolated between measurements.

\subsection{Collision Detection}
In the simulation, rigid objects \comment{(like the robotic finger and the table)} are treated as geometric obstacles.  \comment{When the simulation of the mesh deformation runs---used to model the sponge---the predicted positions of the vertices may indicate that the mesh penetrates an obstacle.  This should be interpreted as a collision between the sponge and an obstacle.  Since the sponge can not penetrate an obstacle, these overlaps must be eliminated. %The physical consequences of this will affect the system predictions.  For example: if the vertices of the mesh overlap the table, the sponge will remain on top of the table, with its material being more compact, which will cause the response force propagated towards the finger to be increased.
}

\comment{In order to model this, we model the finger and table using geometric primitives.  For the table obstacle, any interpenetration by the sponge is solved by pushing vertices of the mesh of the sponge back to the border of the table.  The springs will then propagate the effect to their neighbours.  For the finger obstacle, resolution is in two stages: by vertex and then by edge.}  A vertex will be pushed out to a distance $\epsilon$ of the finger, in the direction of the radius.  For edges, the \comment{standard} algorithm in \alref{alg:point_segment} is used to estimate the shortest distance between the centre of the finger and the edge.  Given this distance, the algorithm in \alref{alg:push_edge} indicates how far and in which direction the edge must be displaced.

\begin{figure}
  \begin{algorithmicieee}{Closest Point on a Segment to Any Other Point}\label{alg:point_segment}
 \STATE Let $a$, $b$ be the ends of the edge $\overline{AB}$, $c$ the external point, and $x$ the closest point to $c$ in the line.
 \STATE Let the vectors $\vec{AB} = b - a$ and $\vec{AC} = c - a$.
 \STATE The projection of $\vec{AC}$ over $\vec{AB}$ is $\|x\| = \dfrac{\vec{AC} \cdot \vec{AB}}{\|\vec{AB}\|}$
 \IF{$\|x\| < 0$}
  \STATE $closest \gets a$
 \ELSE
  \IF{$\|x\| > \|\vec{AB}\| $}
   \STATE $closest \gets b$
  \ELSE
   \STATE $closest \gets a + \|x\| \cdot \dfrac{\vec{AB}}{\|\vec{AB}\|}$
  \ENDIF
 \ENDIF
 \end{algorithmicieee}
 \caption{Algorithm to find the closest point on a segment to any other point.}
\end{figure}

\begin{figure}
  \begin{algorithmicieee}{Push Edge Out}\label{alg:push_edge}
 \STATE Let $a$, $b$ be the ends of the edge $\overline{AB}$, $c$ the centre of the circle.  Use algorithm \ref{alg:point_segment} to find the closest point to the centre in the edge.
 \IF {$\|(closest - c)\| < radius$ (There is an intersection)}
  \STATE $displacement = (radius - \|(closest - c)\| + \epsilon) * \dfrac{(closest - c)}{\|(closest - c)\|}$
  \STATE $ a \gets a + displacement$
  \STATE $ b \gets b + displacement$
 \ENDIF
\end{algorithmicieee}
 \caption{Algorithm to push an edge out of a circular obstacle.}
\end{figure}

\subsection{\comment{Additional} Geometric Constraints}
\comment{While the mass-spring model incorporates constraints based on mesh geometry, these are soft constraints implemented in the form of an energy function. Even with several of these terms incorporated, such as a term for preservation of angles, the mass-spring model can cause triangles of the mesh to overlap one another and be reversed or flattened.  It is thus necessary to employ additional hard geometric constraints to make the mesh internally consistent.  We achieve this through geometric checking subroutines.}  There are two geometric constraints that the mesh is enforced to maintain:

\begin{enumerate}
 \item For a reversed triangle, (one vertex crossed over an edge): the vertex gets pushed more than half the distance between the vertex and the edge in the direction of the ``reversed'' height\footnote{$3/4$ of this distance was used in our implementation.}, while both ends of the edge get pushed the same amount in the opposite direction.  This is enough for most cases, but triangles in the border of the mesh can still overlap each other.

@@

Given the initial shape of the object and a set of parameters, the mass-spring system allows predictions of the deformation of an object for various interactions.  Given the procedures explained in previous sections, there are three ways to calculate the subsequent deformations of the shape:
\begin{enumerate}
 \item \textbf{Collision detection only.} Collision detection is used to determine the compression of the springs around the actuator.  In this case, the neighbouring springs are compressed and their preservation forces propagate the deformation towards their neighbours, and so on.  There is no need to use the force sensor.  It can be equated with observing someone else pushing an object we have touched before.
 \item \textbf{Forces on vertices.} The forces to be applied on the  vertices around the actuator are approximated by the measured reaction forces.  Both the series of positions of the actuator and the reaction forces are necessary.  These forces will be solely responsible for the deformation of the mesh.  If the deformation is insufficient, the actuator may interpenetrate the border of the mesh.  The quality of the simulation is much more sensitive to the values of the elasticity-plasticity constants.  This corresponds to the robot itself pushing the object and feeling the response, while using vision solely to determine the position of the actuator and for evaluation purposes.
 \item \textbf{Collisions and forces.} The previous two approaches can be combined.  Visual constraints (collision detection) and forces are used to determine the displacement of nodes around the actuator.  Again, the springs propagate the deformation to the rest of the mesh.
\end{enumerate}
The mesh as predicted one step ahead is then fed back as the current shape for the next step and the cycle is repeated until the simulation is stopped.

The next stage \comment{is} the learning process \comment{which} consists in identifying sets of parameters that produce simulations similar to the ground truth.  This problem is addressed in the following section.

\section{\comment{Learning to Predict Deformations}}
\label{sec:learning}

@@

\begin{figure}[!t]
\centering
\includegraphics[width=88mm]{arrio2}
\caption{(a) Experimental set up.  A robot finger pushes a block of deformable material sitting on a table.  A camera registers the events of the plane where the main deformations take place.  (b) Evaluation function.  Left top: Mesh over real block of material.  Left bottom: The typical error of the model vs. the ground truth (red) and the typical error of the tracker (blue).  Right top: tracked ground data.  Middle: TP in yellow, FP in blue, FN in red and TN in gray.  Bottom: simulated mesh. }
\label{fig:scene}
\end{figure}

The contour of the deformable object is used to obtain the ground truth for the training phase, which is the area enclosed by the contour. \comment{Since we only require a tracking algorithm which is good enough to evaluate the predictions of the mass-spring system, a very simple one was used.}  Geometrically, the contour is represented by a polygon with hundreds of sides, which can also be called a \textit{linear spline} \comment{\cite{Menet1991}. The vertices of the polygon work as control points which can be moved to adjust the polygon to the shape of the object as it gets deformed.  Even though this implementation is inspired by the theory of active contours, it works with a much simpler and faster, but less accurate, algorithm, instead of minimizing a global energy. This is explained below.}  The tracking of the deformable target in the ground truth has two stages:
\begin{enumerate}
 \item The target object is segmented from the environment.  The Canny algorithm is used to extract the edges of the image.  \comment{  Since the object of interest has a distinctive colour, the hue of the pixels on every edge are used to keep only edges that could belong to it.  If the average hue of the pixels on an edge falls outside a previously calibrated range, the edge is eliminated.}
 \item The internal representation of its contour, the linear snake, is adapted to its new shape.  The algorithm assumes that no occlusions take place and that the contour does not bend over itself.
\end{enumerate}
For the first frame, the snake is initialised as a rectangle around the target. This rectangle fragments itself into smaller pieces. Each new vertex adheres itself to the closest point on an edge, as reached by growing rings of pixels around its current position, up to a certain threshold distance [\fref{fig:scene}(a)].  If there is no edge within that area, the control point remains where it is. This criterion was selected to make the representation more robust against edges that are easier to detect on some frames than in others.  Once the edge reappears, the control point adapts itself again.  Also, the distance between two consecutive control points is kept within the interval $[min\_distance, max\_distance]$ and two adjacent edges do not form angles smaller than $min\_angle$.  See \alref{alg:regularise}.  This allows the snake to adapt very quickly to the complexity of the contour it represents.

\begin{figure}
 \begin{algorithmicieee}{Regularise Linear Spline}\label{alg:regularise}
 \STATE Let $L$ be a linear spline representing the contour of a $2D$ shape, and $L_i$ the $i^{th}$ control point in a cyclic order.
 \FORALL {$L_i$ in $L$}
 \IF {$dist(L_i,L_{i-1}) > max\_distance$}
  \STATE Insert $midpoint(L_i,L_{i-1})$ after $L_i$
 \ELSIF {$dist(L_i,L_{i-1}) < min\_distance$}
  \STATE erase $L_i$
 \ELSIF {$angle(L_{i-1},L_i,L_{i+1}) < min\_angle$}
  \STATE erase $L_i$
 \ENDIF
 \ENDFOR
\end{algorithmicieee}
\caption{Algorithm used to regularise the number of elements of a linear spline, in accordance with the level of detail required.}
\end{figure}

\subsection{Evolutionary Algorithm}
\comment{Since a mass-spring model is not a constitutive model, its parameters are not directly derived from the physical properties of the material, they instead depend on the geometry of the mesh used to model the object. There can thus be several sets of parameters in different regions of the parameter space that produce adequate simulations \cite{Morris2008}. We therefore chose to use an evolutionary algorithm to both perform global search, and local improvement.}

The evolutionary algorithm described in \alref{alg:genetic} was used to search the parameter space of the mass-spring model presented in the previous section.
From the parameters that must be chosen for the simulation to work, some were fixed while others had to be searched for.  They are:
\begin{itemize}
 \item The mass per vertex.  \textit{(For the moment we consider a unitary mass for all vertices.)}
 \item The elasticity constants: for the preservation of length $k_L$, area $k_A$, angles $k_{\varphi}$ and linear damping $k_d$.  \textit{(All springs will have the same values to favour generalisation.)}
 \item The plastic parameters: $yield$, $creep$ and the maximum amount of permanent deformation $max\_\alpha$.  \textit{(All springs will have the same values to favour generalisation.)}
 \item The integration time step $\Delta t$.  \textit{(It has been set to 0.1 seconds, even though values as small as 0.01 were tried as well.)}
 \item An upper limit for the magnitude of the total estimated forces per vertex $max\_Force$.
 \item The minimum area before a triangle is considered \textit{flat}.  (It has been set to $2.0$ units.)
\end{itemize}
To evaluate the suitability of each set of parameters, the simulation is compared with the ground truth tracked from real objects.

\begin{figure}
  \begin{algorithmicieee}{Evolutionary Search}\label{alg:genetic}
 \STATE Let $N$ be the number of sets of parameters to be evaluated per generation and $M$ the number of generations.
 \STATE Let $\alpha + \beta + \gamma + \zeta = N$ denote four portions of the $N$ elements\comment{, with: \\
 $\alpha \leftarrow$ number of elements that survive for the next generation as they are, \\
 $\beta \leftarrow$ mutated individuals, \\
 $\gamma \leftarrow$ crossed over, and \\
 $\zeta \leftarrow$ new individuals. \\
 which will be chosen as follows:}
 \STATE Let $\sigma_{max} > \sigma_{min}$ both be standard deviations of a Gaussian
 \STATE Let $\Delta\sigma \gets (\sigma_{max} - \sigma_{min})/M$.
 \FOR {$i=1 \to M$}
 \STATE Run $N$ simulations and compare prediction with ground truth (real object), by comparing the area occupied by the real object with the area occupied by the simulated mesh [\fref{fig:scene}(b)].
 \STATE Keep the best $\alpha$ from all simulations.
 \STATE $\sigma \gets \sigma - \Delta\sigma$
 \STATE Generate $\beta$ by adding random Gaussian noise, with standard deviation $\sigma$, to the best candidates.
 \STATE Generate $\gamma$ by selecting values of the parameters from two different sets at random, after eliminating the $\epsilon$ worst.
 \FOR {$\zeta$ elements}
   \STATE Generate completely new sample with:
   \FOR {every parameter}
     \STATE $n \leftarrow randomNumber \in [0,1)$
     \STATE $min \leftarrow \log(minValue)$
     \STATE $max \leftarrow \log(maxValue)$
     \STATE $param \leftarrow e^{(n (max - min) + min)}$
   \ENDFOR
 \ENDFOR
 \STATE $i \gets i + 1$
 \ENDFOR
\end{algorithmicieee}
 \caption{Evolutionary Search for Automatic Calibration}
\end{figure}

@@

To evaluate the quality of the predicted deformations, the pixels inside the tracked contour are used as ground truth.  These are  compared against those inside the mesh of the simulation.  In machine learning terms, these pixels can be divided into four classes: the \textit{true positives $TP$}, where the material is present, and the simulation predicted it would be present; the \textit{true negatives $TN$}, where the absence of material is in accordance with the prediction; the \textit{false positives $FP$}, where the prediction said there would be material, but there is not; and the \textit{false negatives $FN$}, where there is material, but the prediction said there wouldn't be [\fref{fig:scene}(b)].  Given that maintaining a record of the true negatives (the empty/background space around the material) is mostly irrelevant and highly inefficient, the score for a whole video was derived from the following quantities:
\begin{align}
 Precision &= \frac{TP}{TP+FP}, \\
 Recall &= \frac{TP}{TP+FN}
\end{align}
From which the harmonic average $F$ is derived:
\begin{align}
 F = \frac{2}{\frac{1}{Precision} + \frac{1}{Recall}} = \frac{1}{1 + \frac{FN + FP}{2TP}}
\end{align}

This measurement \comment{is know as the F-score and} is used in the computer vision community to evaluate segmentation algorithms.  $F=1$ means that there were no false predictions ($FN+FP = 0$).  $F \rightarrow 0$\footnote{This is read: the value of $F$ goes to zero or $F$ tends to zero.}, if there are no good predictions ($TP\rightarrow0$) or if there are too many errors ($FN+FP \rightarrow \infty$).  $F$ is calculated for every frame and the score for the video \comment{ at frame $i$} is the mean over all the frames so far, $\mu(F)$:
\comment{\begin{align}
 \mu(F) = \sum_{t=0}^{t=i} F(frame=t)
\end{align}}
If the predictions for all the frames were perfect, $\mu(F)$ would be equal to 1; the closer to zero, the more errors there were.  The simulation stops sometimes, if a model has become unstable; from that frame on, the value of $F$ is zero.  The model with the highest final value of $\mu(F)$ is considered the best.

\section{\comment{Learning to predict forces}}
%We used two methods to predict the forces.  The first one consists in measuring the total force acting on the vertex of the mesh closest to the actuator, according to the mass-spring model, after solving detected collisions.  This force is feed as input for the following simulation step.
\label{sec:forces}

Additionally and independently from the mass-spring system,  it is possible to learn to make predictions about the expected reaction force of the material, by using information from the stress-strain diagram obtained from the \comment{same} training data set \comment{as the mass-spring model.  This diagram is obtained from the video recorded by the visual system of the robot and the forces recorded by a sensor located at the wrist.  When the readings of the force sensor start to increase and the figer collides with the block of material, the point of zero deformation is identified.  From here on, the displacement of the finger is considered equivalent to the amount of deformation in the material.  Even though this method is not as rigorous as an engineering test, the estimation is good enough for robot manipulation of common materials.}
